基于Lab色彩空间的太阳黑子手绘图背景提取方法

### 2017-01-20 paper/paper.tex
修改摘要，添加最近质心分类器

\textbf{摘要: }太阳黑子手绘图是记录太阳黑子活动的传统方式，手绘图的数字化有利于数据的长期保存和研究，而背景提取是数字化过程中的一个关键问题。针对该问题，本文提出了一种基于Lab色彩空间的快速背景提取方法。该方法在Lab色彩空间中，对手绘图的$a$、$b$色彩分量进行两类K-means聚类。聚类结果表明，背景与手绘部分准确的分割为两类。对于大批量手绘图背景提取的情况，为加快提取速度，仅抽取部分手绘图进行聚类，再利用样本的聚类结果直接对其他手绘图进行背景提取。实验结果表明，此方法在保证提取准确性的前提下，极大地加快了提取速度。

\textbf{关键词: }背景提取；太阳黑子；Lab色彩空间；K-means聚类

\textbf{Abstract: }Sunspots drawings is a traditional way of recording sunspots activities. Digitization of the drawings is beneficial to long-term storage and research of the data. Background extraction is an important step in the process of digitization. For this problem, we develope a rapid background extraction method based on Lab color space. In this method, we perform two-class K-means cluster for the $a$, $b$ component of Lab color space on the drawings. The cluster results show that the background and the handdrawing part are clustered into two class accurately. For mass drawings background extraction, we only select a part of drawings to perform the aformentioned cluster, then apply their result on other drawings directly. The experimental results show that this method not only enormously speed up the time of processing, but also guaranteed the accuracy.

\textbf{Keywords: }background extraction；sunspot；CIELAB color space；k-means clustering

>>>>

\textbf{摘要: }太阳黑子手绘图是记录太阳黑子活动的传统方式，手绘图的数字化有利于数据的长期保存和研究，而背景提取是数字化过程中的一个关键问题。针对该问题，本文提出了一种基于Lab色彩空间的快速背景提取方法。该方法在Lab色彩空间中，对手绘图的$a$、$b$色彩分量进行两类K-means聚类。聚类结果表明，背景与手绘部分准确的分割为两类。对于大批量手绘图背景提取的情况，为加快提取速度，仅抽取部分手绘图进行聚类，再利用样本的聚类结果用最近质心分类器对其他手绘图进行背景提取。实验结果表明，此方法在保证提取准确性的前提下，极大地加快了提取速度。

\textbf{关键词: }背景提取；太阳黑子；Lab色彩空间；K-means聚类；最近质心分类器

\textbf{Abstract: }Sunspots drawings is a traditional way of recording sunspots activities. Digitization of the drawings is beneficial to long-term storage and research of the data. Background extraction is an important step in the process of digitization. For this problem, we develope a rapid background extraction method based on Lab color space. In this method, we perform two-class K-means cluster for the $a$, $b$ component of Lab color space on the drawings. The cluster results show that the background and the handdrawing part are clustered into two class accurately. For mass drawings background extraction, we only select a part of drawings to perform the aformentioned cluster, then apply nearest centroid classifier on other drawings directly. The experimental results show that this method not only enormously speed up the time of processing, but also guaranteed the accuracy.

\textbf{Keywords: }background extraction；sunspot；CIELAB color space；k-means clustering；nearest centroid classifier
